---
title: "Seagrass Thermal Tolerance"
author: "Martina Burgo"
date: today
date-format: "DD/MM/YYYY"
format: 
  html:
    ## Format
    theme: default
    ## Table of contents
    toc: true
    toc-float: true
    ## Numbering
    number-sections: true
    number-depth: 3
    ## Layout
    fig-caption-location: "bottom"
    fig-align: "center"
    fig-dpi: 72
    tbl-cap-location: top
    ## Code
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
    code-line-numbers: true
    ## Execution
    execute:
      echo: true
      cache: true
    ## Rendering
    embed-resources: true
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
  tbl-title: '**Table**'
  tbl-labels: arabic
engine: knitr
bibliography: ../docs/resources/references.bib
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
classoption: a4paper
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(cache.lazy = FALSE,
                      tidy = "styler")
options(tinytex.engine = "xelatex")
```

# Intro

This dataset contains the data gathered from the systematic search and the climate variables calculated as summarised

## Climate variables

Data on mean monthly sea surface temperatures for the analysis was downloaded from the National Oceanic and Atmospheric Administration (https://psl.noaa.gov/data/gridded/data.cobe2.html). The variable used is 'sst.mon.mean.nc'. To investigate how seagrass thermal tolerance varies between species and populations, the following variables were created for each seagrass population:

| Variable                                    | At population-level                       | At species-level                                                  |
|---------------------|--------------------|-------------------------------|
| MAT (mean annual temperature)               | Average MAT across years for one location | Average MAT across years averaged across the species distribution |
| AV (annual variation)                       | Average AV across years for one location  | Max AV across years averaged across the species distribution      |
| MTWA (maximum temperature of warmest moth)  | Max MTWA across years for one location    | Max MTWA across years averaged across the species distribution    |
| Difference (Experimental temperature - MAT) |                                           |                                                                   |
| Magnitude (Difference/AV)                   |                                           |                                                                   |


# Preparations

## Load the necessary libraries

```{r}
#| label: libraries
#| output: false
#| eval: true
#| warning: false
#| message: false
#| cache: false

library(tidyverse)   #for data wrangling
library(stringr)
library(car)
library(brms)
library(patchwork)
library(corrplot)
library(loo)
library(tidybayes)
library(DHARMa)     #for residual diagnostics
library(rstan)      #for interfacing with STAN
library(emmeans)
library(ggridges)   #for ridge plots 
library(raster)

source('../R/functions.R')
```

## Read in the data
```{r}
#| label: loadModels
#| eval: false

load(file = '../data/modelled/Model6Beta.RData')
load(file = '../data/modelled/Model7Beta.RData')
load(file = '../data/modelled/Model8Beta.RData')

pred_1w_stack <- raster::stack('../data/processed/pred_1w_stack.grd')
pred_2w_stack <- raster::stack('../data/processed/pred_2w_stack.grd')
```


# Exploratory analysis

Visual representation of the data:

```{r}
seagrass |> 
  ggplot(aes(x = log(Time), y = Temperature, color = Survival)) +
  geom_point(size = 3) + scale_x_continuous('Time (log)') + 
  scale_color_viridis_c(option = 'C') + theme_classic()
```

```{r}
seagrass |>
  ggplot(aes(y = Survival, x = difference_species, color = av_species)) +
  geom_point(size = 3) + theme_classic() + scale_x_continuous('Temp difference (T-MAT)')
```

And plotting the potential predictors to check if they're correlated:

::: panel-tableset
## Scatterplot

```{r}
#| label: scatterPlot
scatterplotMatrix(~Survival+log(Time)+Temperature+mat_population+av_population+mtwa_population+difference_population + av_species + mat_species + mtwa_species + difference_species, data = seagrass,diagonal = list(method = 'boxplot'))
```

## Correlation plot

```{r}
#| label: corrPlot
seagrass |> 
  dplyr::select(Time, Temperature, mat_population, av_population, mtwa_population, difference_population, av_species, mat_species, mtwa_species, difference_species) |> 
  cor() |> 
  corrplot()
```
:::

# Results

## M8: Basic model

### Effect sizes

```{r}
seagrass.brm8 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |> #to go from log scale to odds ratio
  dplyr::select(matches("^b_.*|^sd_.*")) |> 
  summarise_draws(median, HDInterval::hdi, rhat, length, ess_bulk, ess_tail,
                  Pl = ~ mean(.x < 1),
                  Pg = ~ mean(.x > 1)) |>
  as_tibble()
```

```{r}
1.05/(1+1.05)
#the intercept is relative to the average of all other predictors
```

```{r}
100*(0.71-1) #for each unit increase of time, the rate of survival declines by 29% 
```

```{r}
100*(0.48-1) #for each unit increase of temperature, the rate of survival declines by 52% 
```

```{r}
100*(1.44-1) #on average, Tropical survival is 44% higher
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

seagrass.brm8 |> 
  gather_draws(`^b_.*`, regex = TRUE) |>
  filter(.variable != 'b_Intercept') |> 
  ggplot() +
  geom_density_ridges_gradient(aes(x = .value,
                                   y = .variable,
                                   fill = stat(x)),
                               alpha = 0.4, colour = 'white',
                               quantile_lines = TRUE,
                               quantiles = c(0.025, 0.975)) +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  scale_x_continuous(name = '') +
  scale_y_discrete(name = '', labels = c('Time*', 'Time:Temperature', 'Temperature*', 'Type')) +
  scale_fill_viridis_c(option = 'D') + theme_classic() +
  theme(text = element_text(colour = 'black'), 
        legend.position = 'none', 
        axis.text = element_text(size = rel(1.5)))
```

### Survival ~ Temp + Time

```{r}
newdata <- with(seagrass,
                list(Temperature = seq(from = 1,
                                       to = max (Temperature),
                                       len = 20),
                            Time = seq(from = 1,
                                       to = 336,
                                       len = 14))) #two weeks of heat stress, at 1-day intervals
fit <- seagrass.brm8 |> 
  emmeans(~Temperature|Time, at = newdata) |>
  gather_emmeans_draws() |>
  mutate(.value = plogis(.value)) |>
  group_by(Temperature, Time) |>
  summarise(median = median(.value)) |>
  as.data.frame()
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

fit |> 
  ggplot(aes(y = Temperature, x = Time, fill = median)) + 
  geom_tile() +
  scale_x_continuous(name = 'Time (days)', 
                     n.breaks = 14, 
                     labels = c((seq(24, 336, len = 14)/24))) + #adjusts the labels to show days instead of hours
  scale_y_continuous(expression("Temp difference ("*~degree*C*")"),
                     n.breaks = 10) +
  scale_fill_viridis_c(option = 'C', name = 'Survival') + theme_classic() +
  theme(text = element_text(colour = 'black'), 
        axis.text = element_text(size = rel(1.2)),
        axis.title = element_text(size = rel(1.5)))
```

## M6:Population-level model

### Effect sizes

```{r}
seagrass.brm6 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |> #to go from log scale to odds ratio
  dplyr::select(matches("^b_.*|^sd_.*")) |> 
  summarise_draws(median, HDInterval::hdi, rhat, length, ess_bulk, ess_tail,
                  Pl = ~ mean(.x < 1),
                  Pg = ~ mean(.x > 1)) |>
  as_tibble()
```

```{r}
1.2/(1+1.2)
#the intercept is relative to the average of all other predictors
```

```{r}
100*(0.69-1) #for each unit increase of time, the rate of survival declines by 31% 
```

```{r}
100*(0.56-1) #for each unit increase of diff_pop, the rate of survival declines by 44% 
```

```{r}
100*(1.33-1) #for each unit increase of av_pop, the rate of survival increase by 33%
```

```{r}
100*(0.71-1) #for each unit increase of mtwa_pop, the rate of survival declines by 29% 
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

seagrass.brm6 |> gather_draws(`^b_.*`, regex = TRUE) |>
  filter(.variable != 'b_Intercept') |>
  ggplot() +
  geom_density_ridges_gradient(aes(x = .value,
                                   y = .variable,
                                   fill = stat(x)),
                               alpha = 0.4, colour = 'white',
                               quantile_lines = TRUE,
                               quantiles = c(0.025, 0.975)) +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  scale_x_continuous(name = '') +
  scale_y_discrete(name = '', labels = c('AV*', 'Difference*', 'Time*', 'Time:Difference', 'MTWA*')) +
  scale_fill_viridis_c(option = 'D') + theme_classic() +
  theme(text = element_text(colour = 'black'), 
        legend.position = 'none', 
        axis.text = element_text(size = rel(1.5)))
```

### Survival ~ Diff + AV

```{r}
newdata2 <- with(seagrass,
                list(difference_population = seq(from = 1,
                                                        to = max (difference_population),
                                                        len = 20),
                     av_population = c(quantile(av_population)[2] |> round(0),
                                       quantile(av_population)[5] |> round(0))))
fit2 <- seagrass.brm6 |> 
  emmeans(~difference_population|av_population, at = newdata2) |>
  gather_emmeans_draws() |>
  mutate(.value = plogis(.value)) |>
  group_by(difference_population, av_population) |>
  summarise(median_hdci(.value)) |>
  as.data.frame()
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

fit2 |> 
  mutate(av_population = factor(av_population)) |>
  ggplot(aes(x = difference_population,
             y = y, 
             color = av_population)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = av_population), alpha = 0.2) +
  scale_color_discrete(guide = 'none') +
  scale_fill_discrete('AV') +
  scale_x_continuous(expression("Temp difference ("*~degree*C*")"),
                     n.breaks = 10) +
  scale_y_continuous('Survival', limits = c(0,1)) +
  theme_classic()  +
  theme(text = element_text(colour = 'black'), 
        axis.text = element_text(size = rel(1.2)),
        axis.title = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.5)))
```

At what range of temperature difference does higher AV results in survival increase of at least 25%?

```{r}
seagrass.brm6 |> 
  emmeans(~difference_population|av_population, at = newdata2, type = 'response') |> 
  gather_emmeans_draws() |> 
  mutate(Sur_diff = plogis(.value)) |>
  group_by(.draw, difference_population) |>
  summarise(Surv_diff = diff(Sur_diff)) |>
  ungroup() |>
  filter(Surv_diff >= 0.25) |>
  summarise(median_hdci(difference_population),
            Pl = mean(difference_population < 1),
            Pg = mean(difference_population > 1))
```

At what temp difference is the survival improvement highest between AV?

```{r}
seagrass.brm6 |> 
  emmeans(~difference_population|av_population, at = newdata2, type = 'response') |> 
  gather_emmeans_draws() |> 
  mutate(Sur_diff = plogis(.value)) |>
  group_by(.draw, difference_population) |>
  summarise(Sur_diff = diff(Sur_diff)) |>
  ungroup() |>
  group_by(.draw) |>
  filter(difference_population == difference_population[which.max(Sur_diff)]) |>
  ungroup() |>
  summarise(median_hdci(difference_population),
            Pl = mean(difference_population < 1),
            Pg = mean(difference_population > 1))
```

### Survival ~ Diff + MTWA

```{r}
newdata3 <- with(seagrass,
                list(difference_population = seq(from = 1,
                                                        to = max (difference_population),
                                                        len = 20),
                     mtwa_population = c(quantile(mtwa_population)[2] |> round(0),
                                         quantile(mtwa_population)[5] |> round(0))))
fit3 <- seagrass.brm6 |> 
  emmeans(~difference_population|mtwa_population, at = newdata3) |>
  gather_emmeans_draws() |>
  mutate(.value = plogis(.value)) |>
  group_by(difference_population, mtwa_population) |>
  summarise(median_hdci(.value)) |>
  as.data.frame()
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

fit3 |> 
  mutate(mtwa_population = factor(mtwa_population)) |>
  ggplot(aes(x = difference_population,
             y = y, 
             color = mtwa_population)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = mtwa_population), alpha = 0.2) +
  scale_color_discrete(guide = 'none') +
  scale_fill_discrete('MTWA') +
  scale_x_continuous(expression("Temp difference ("*~degree*C*")"),
                     n.breaks = 10) +
  scale_y_continuous('Survival', limits = c(0,1)) +
  theme_classic()  +
  theme(text = element_text(colour = 'black'), 
        axis.text = element_text(size = rel(1.2)),
        axis.title = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.5)))
```

### LD50

For Temp diff

```{r}
seagrass.brm6 |> 
  emmeans(~difference_population, 
          at = with(seagrass, 
                    list(difference_population = seq(from = 1,
                                                     to = max (difference_population),
                                                     len = 20))), 
                    type = 'response') |> 
  gather_emmeans_draws() |> 
  mutate(.value = plogis(.value)) |>
  filter(.value >= 0.49 & .value <= 0.51) |>
  ungroup() |>
  summarise(median_hdci(difference_population),
            Pl = mean(difference_population < 1),
            Pg = mean(difference_population > 1))
```

For Time (in hours) at three temp diff

```{r}
seagrass.brm6 |> 
  emmeans(~Time|difference_population, 
          at = with(seagrass, 
                    list(Time = seq(from = 1,
                                    to = max(Time),
                                    len = 50),
                         difference_population = c(5, 10, 15))), 
                    type = 'response') |> 
  gather_emmeans_draws() |>
  mutate(.value = plogis(.value)) |>
  filter(.value >= 0.49 & .value <= 0.51) |>
  ungroup() |>
  group_by(difference_population) |>
  summarise(median_hdci(Time),
            Pl = mean(Time < 1),
            Pg = mean(Time > 1))
```

## M7: Species-level model

### Effect sizes

```{r}
seagrass.brm7 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |> #to go from log scale to odds ratio
  dplyr::select(matches("^b_.*|^sd_.*")) |> 
  summarise_draws(median, HDInterval::hdi, rhat, length, ess_bulk, ess_tail,
                  Pl = ~ mean(.x < 1),
                  Pg = ~ mean(.x > 1)) |>
  as_tibble()
```

```{r}
1.17/(1+1.16)
#the intercept is relative to the average of all other predictors
```

```{r}
100*(0.78-1) #for each unit increase of time, the rate of survival declines by 22% 
```

```{r}
100*(0.59-1) #for each unit increase of diff_spp, the rate of survival declines by 41% 
```

```{r}
100*(0.7-1) #for each unit increase of mtwa_spp, the rate of survival declines by 30% 
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

seagrass.brm7 |> gather_draws(`^b_.*`, regex = TRUE) |>
  filter(.variable != 'b_Intercept') |>
  ggplot() +
  geom_density_ridges_gradient(aes(x = .value,
                                   y = .variable,
                                   fill = stat(x)),
                               alpha = 0.4, colour = 'white',
                               quantile_lines = TRUE,
                               quantiles = c(0.025, 0.975)) +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  scale_x_continuous(name = '') +
  scale_y_discrete(name = '', labels = c('AV', 'Difference*', 'Time*', 'Time:Diff', 'MTWA*')) +
  scale_fill_viridis_c(option = 'D') + theme_classic() +
  theme(text = element_text(colour = 'black'), 
        legend.position = 'none', 
        axis.text = element_text(size = rel(1.5)))
```

### Survival ~ Diff + MTWA

```{r}
newdata4 <- with(seagrass,
                list(difference_species = seq(from = 1,
                                                        to = max (difference_species),
                                                        len = 20),
                     mtwa_species = c(quantile(mtwa_species)[2] |> round(0),
                                         quantile(mtwa_species)[5] |> round(0))))
fit4 <- seagrass.brm7 |> 
  emmeans(~difference_species|mtwa_species, at = newdata4) |>
  gather_emmeans_draws() |>
  mutate(.value = plogis(.value)) |>
  group_by(difference_species, mtwa_species) |>
  summarise(median_hdci(.value)) |>
  as.data.frame()
```

```{r}
#| fig-width: 31.49
#| fig-height: 19.68

fit4 |> 
  mutate(mtwa_species = factor(mtwa_species)) |>
  ggplot(aes(x = difference_species,
             y = y, 
             color = mtwa_species)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = mtwa_species), alpha = 0.2) +
  scale_color_discrete(guide = 'none') +
  scale_fill_discrete('MTWA') +
  scale_x_continuous(expression("Temp difference ("*~degree*C*")"),
                     n.breaks = 10) +
  scale_y_continuous('Survival', limits = c(0,1)) +
  theme_classic()  +
  theme(text = element_text(colour = 'black'), 
        axis.text = element_text(size = rel(1.2)),
        axis.title = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.5)))
```

### LD50

For Temp diff

```{r}
seagrass.brm7 |> 
  emmeans(~difference_species, 
          at = with(seagrass, 
                    list(difference_species = seq(from = 1,
                                                     to = max (difference_species),
                                                     len = 20))), 
                    type = 'response') |> 
  gather_emmeans_draws() |> 
  mutate(.value = plogis(.value)) |>
  filter(.value >= 0.49 & .value <= 0.51) |>
  ungroup() |>
  summarise(median_hdci(difference_species),
            Pl = mean(difference_species < 1),
            Pg = mean(difference_species > 1))
```

For Time (in hours) at three temp diff

```{r}
seagrass.brm7 |> 
  emmeans(~Time|difference_species, 
          at = with(seagrass, 
                    list(Time = seq(from = 1,
                                    to = max(Time),
                                    len = 50),
                         difference_species = c(5, 10, 15))), 
                    type = 'response') |> 
  gather_emmeans_draws() |>
  mutate(.value = plogis(.value)) |>
  filter(.value >= 0.49 & .value <= 0.51) |>
  ungroup() |>
  group_by(difference_species) |>
  summarise(median_hdci(Time),
            Pl = mean(Time < 1),
            Pg = mean(Time > 1))
```

# Compare models

```{r}
loo::loo_compare(rstan::loo(seagrass.brm6),
                 rstan::loo(seagrass.brm7))
```

# Identify vulnerable communities

```{r}
as.data.frame(table(as.matrix(pred_1w_stack[[1]]))) |>
  mutate(Percent = Freq/sum(Freq) * 100) |>
  head()
```

# Supplementary materials

# Session Info

```{r}
#| label: sessionInfo
sessionInfo()
```
